---
title: "Analysis of Montreal breakins data 2015-16"
author: "Quan Nguyen"
date: "May 18, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Get data

Data source:

http://donnees.ville.montreal.qc.ca/dataset/actes-criminels

```{r}
library(data.table)
library(dplyr)
library(dygraphs)
library(ggplot2)
DT = fread(input = "donneesouvertes-citoyens.csv", header = T, sep=",", stringsAsFactors = T)
DT
DT %>% summary()
str(DT)
```

# Plot over time

```{r}
library(zoo)
# Add a few date attributes for timeseries processing
DT$mydate = as.Date(as.character(DT$DATE), tz="EST", format="%Y%m%d")
DT$year = as.Date(cut(DT$mydate, breaks="year"))
DT$month = as.Date(cut(DT$mydate, breaks="month"))
DT$week = as.Date(cut(DT$mydate, breaks="week"))
DT$weekday = as.POSIXlt(DT$mydate)$wday

# plot of daily breakins split by period (day, evening, night)
ggplot(DT, aes(x=mydate, color=QUART)) + geom_freqpoly()
ggplot(DT, aes(x=mydate)) + facet_grid(QUART ~ .) + geom_freqpoly()
# plot number of breakins per week
plot(table(DT$week), main="By Week", ylab="Breakins", type='l')
# Plot nunber of breakins per month split by period
ggplot(DT, aes(x=month, fill=QUART)) + geom_bar()
# By weekday (0: Sunday)
table(DT$weekday);
ggplot(DT, aes(x=weekday, fill=QUART)) + geom_bar()
# By Quart
table(DT$QUART); 
ggplot(DT, aes(x=QUART, fill=QUART)) + geom_bar()
```

# Create daily aggregate

``` {r}
DT_daily = count(DT, mydate)
```

# Convert to ts() series

```{r}
# frequency=365 will give proper x-axis tic values
# however will not be accepted by stl() because of
# series is not periodic or has less than two periods
DT_daily.ts = ts(
  DT_daily$n,
  frequency=30 #, start=c(2015,0.1)
)
plot(DT_daily.ts)
# timeseries decomposition
DT_daily.stl = stl(DT_daily.ts, s.window=7)
plot(DT_daily.stl)

```



``` {r}
# Convert to zoo series

DT_daily.z = zoo(
  DT_daily$n,
  order.by=DT_daily$mydate,
  frequency=7
)
```

``` {r eval=F, echo=F}
sub_text = range(index(DT_daily.z))
#plot(DT_daily.z, type="s")
dygraph(DT_daily.z)

DT_daily.z.decompose = decompose(ts(DT_daily.z,frequency=30), "additive")
plot(DT_daily.z.decompose, xaxt='n')
title(sub=paste(sub_text,collapse="  to  "))
```

# Arima forecast

```{r}
library(forecast)
library(ggplot2)
DT_daily.z.forecast =
  forecast.Arima(
    auto.arima(
      DT_daily.ts
    ), 
    h=60
  )
autoplot(DT_daily.z.forecast)

# prediction of the trendline

DT_daily.ts.forecast =
  forecast.Arima(
    auto.arima(
      DT_daily.stl$time.series[,2]
    ), 
    h=60
  )
autoplot(DT_daily.ts.forecast)
```

# Anomaly detection

```{r}
library(AnomalyDetection)
myts = as.data.frame(
  cbind(
    as.POSIXct(index(DT_daily.z)),
    coredata(DT_daily.z)
  )
)
data_anomaly = AnomalyDetectionTs(myts, max_anoms=0.01, direction="pos", plot=F, e_value = T, na.rm = T)
# No anomaly detected as NULL result returned
data_anomaly
data_anomaly$plot
```

  
# EDA

```{r}
# Breakins by PDQ (Police de quartier) sorted
DT %>% select(PDQ) %>% table %>% sort(decreasing=T) 
DT %>% group_by(PDQ) %>% summarise(n = n()) %>% mutate(freq = n / sum(n)) %>% arrange(desc(freq)) %>% select(PDQ, freq)

DT %>% select(PDQ, QUART) %>% table
# columns without/with zero values
colSums(DT == 0)
colSums(DT != 0)
DT_goodXY = DT %>% filter(X != 0); DT_goodXY
library(ggplot2)
# ggplot(DT_goodXY, aes(y=LAT, x=LONG, col=DT_goodXY$QUART)) + geom_point()
```

# Put it on the map

```{r fig.width=12, fig.height=12}
library(maptools)
# needs all the files shipped in addition to .shp files
mtl_admin_shp = readShapeSpatial("data/LIMADMIN.shp")
mtl_admin_poly = readShapePoly("data/LIMADMIN") 

mtl_admin_points = fortify(mtl_admin_shp)
centroids.df <- as.data.frame(coordinates(mtl_admin_poly))
names(centroids.df) <- c("long", "lat")
centroids.df$id = as.character(mtl_admin_poly$NOM)

# in ggplot over ggmap
library(ggmap)
montreal12 <- get_map(location = "montreal", zoom=12)
gc3 = geocode("Pont Viau, Quebec", source="google")

center <- as.numeric(gc3)
montreal12 <- get_googlemap(location = "montreal", zoom=11, center=center)

ggmap(montreal12) +
  coord_equal() +
  geom_point(data=DT_goodXY, aes(y=LAT, x=LONG, col=DT_goodXY$QUART)
  )
```

# Conclusion

* Breakins occur motly in daytime or evening, much less during the night
* Breaskins are highest in Oct-Dec
* Breaksins are highest on Monday
* The data is just for 15 months so it's not enough to get a feel for yearly trend
* Police de quartier stations 38, 26, 23 see the most number of breakins (10%, 6.7% and 6.3% respectively)
* Note that the map only shows areas that are part of the city of Montreal only. There doesn't seem any part that has higher infractions
